Can the symmetric Fermi and eROSITA bubbles be produced by tilted jets: a proof-of-concept study

### 2022-03-14 mnras_guide.tex
update

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. H. Tseng et al.]{\large
Po-Hsun Tseng$^{1}$%
\thanks{\href{mailto:zengbs@gmail.com}{zengbs@gmail.com}}%
\orcidlink{https://orcid.org/0000-0002-1868-0660}%
\quad H.-Y. Karen Yang$^{2,3,4,5}$%
\thanks{\href{mailto:hyang@phys.nthu.edu.tw}{hyang@phys.nthu.edu.tw}}%
\orcidlink{https://orcid.org/0000-0003-3269-4660}%
\quad Hsi-Yu Schive$^{1,5,6,7}$%
\thanks{\href{mailto:hyschive@phys.ntu.edu.tw}{hyschive@phys.ntu.edu.tw}}%
\orcidlink{https://orcid.org/0000-0002-1249-279X}%
\quad Chun-Yen Chen$^{8}$
\quad Tzihong Chiueh$^{1,5,6}$%
\thanks{\href{mailto:chiuehth@phys.ntu.edu.tw}{chiuehth@phys.ntu.edu.tw}}%
\orcidlink{https://orcid.org/0000-0003-2654-8763}%
\\\\
$^{1}$Institute of Astrophysics, National Taiwan University, Taipei 10617, Taiwan\\
$^{2}$Institute of Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan\\
$^{3}$Department of Physics, National Tsing Hua University, Hsinchu 30013, Taiwan\\
$^{4}$Center for Informatics and Computation in Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan\\
$^{5}$Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan\\
$^{6}$Department of Physics, National Taiwan University, Taipei 10617, Taiwan\\
$^{7}$Center for Theoretical Physics, National Taiwan University, Taipei 10617, Taiwan\\
$^{8}$Institute of physics, National Taiwan University, Taipei 10617, Taiwan\\
}

@@

% Abstract of the paper
\begin{abstract}
The \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ on either side of the Galactic center (GC).\
The recent discovery of giant eROSITA bubbles also shows a symmetrically enhanced X-ray\
emission from the shells around the eROSITA bubbles.\
To this end, previous simulation-based works based on active galactic nucleus (AGN) jets model\
assumed that the jets are vertical to the Galactic disk;\
however, there is a lack of preferred orientation of jets with respect to the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic (SRHD) simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium (ISM) disk is a substantial\
component for the symmetric \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton (IC) scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
an oblique AGN jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to\
the rotation axis of the Galatic plane (disc normal hereafter).\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a\
hot ($\sim2$ keV), turbulent, and pressure-balanced plasma,\
which we interpret as the \textit{Fermi} bubbles.\
Inside the simulated \textit{Fermi} bubbles,\
there are two additional innermost bubbles we have not detected yet.\
The innermost bubbles are essentially a reverse shock permanently associated with the disk.\
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between the simulations and observations\
, either from the point of view of their morphology or multiwavelength features,\
has thus released a caveat:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

Their symmetry about the GC suggests that they originate\
from powerful energy injections from the GC, possibly related to nuclear star formation\
\citep{PhysRevLett.106.101102,Carretti2013}
or past AGN activity \citep{Guo2012,Yang2017}.

@@

 Followed by the forward shock is a turbulent and hot plasma\
 in pressure balance with the external medium.\
 As shown in the temperature or density slices of the fiducial run,\
 the full extent of the turbulent plasma is 9 kpc, fairly close to\
 that of the observed \textit{Fermi} bubbles \citep{Su2010}.
 Also, the temperature of the plasma is as high as 2 keV,\
 comparable to few keV inside the \textit{Fermi} bubbles\
 by observing X-ray absorption lines through the hot\
 gaseous halo along many different sight lines in the sky \citep{Miller_2013}.\
 The general agreement of extent and temperature\
 suggests that the \textit{Fermi} bubbles\
 is essentially a turbulence rather than a smooth shock \citep{Zhang2020}.

@@

 Moreover, the close-up view (right column in Fig. \ref{fig__profile})\
 of profiles and slices (Fig. \ref{fig__innerbubbles})\
 demonstrates that the pressure rises dramatically outward across\
 the edge of the innermost bubbles,\
 indicating that the innermost bubbles are an expanding reverse shock.\
 The high density upstream of the reverse shock requires\
 an even higher density downstream.\
 Continuing outward, the higher density must match with low density further downstream;\
 thus there exists a dense shell.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. H. Tseng et al.]{\large
Po-Hsun Tseng$^{1}$%
\thanks{\href{mailto:zengbs@gmail.com}{zengbs@gmail.com}}%
\orcidlink{https://orcid.org/0000-0002-1868-0660}%
\quad H.-Y. Karen Yang$^{2,3,4,5}$%
\thanks{\href{mailto:hyang@phys.nthu.edu.tw}{hyang@phys.nthu.edu.tw}}%
\orcidlink{https://orcid.org/0000-0003-3269-4660}%
\quad Hsi-Yu Schive$^{1,5,6,7}$%
\thanks{\href{mailto:hyschive@phys.ntu.edu.tw}{hyschive@phys.ntu.edu.tw}}%
\orcidlink{https://orcid.org/0000-0002-1249-279X}%
\quad Chun-Yen Chen$^{8}, $
\quad Tzihong Chiueh$^{1,5,6}$%
\thanks{\href{mailto:chiuehth@phys.ntu.edu.tw}{chiuehth@phys.ntu.edu.tw}}%
\orcidlink{https://orcid.org/0000-0003-2654-8763}%
\\\\
$^{1}$Institute of Astrophysics, National Taiwan University, Taipei 10617, Taiwan\\
$^{2}$Institute of Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan\\
$^{3}$Department of Physics, National Tsing Hua University, Hsinchu 30013, Taiwan\\
$^{4}$Center for Informatics and Computation in Astronomy, National Tsing Hua University, Hsinchu 30013, Taiwan\\
$^{5}$Physics Division, National Center for Theoretical Sciences, Taipei 10617, Taiwan\\
$^{6}$Department of Physics, National Taiwan University, Taipei 10617, Taiwan\\
$^{7}$Center for Theoretical Physics, National Taiwan University, Taipei 10617, Taiwan\\
$^{8}$Institute of physics, National Taiwan University, Taipei 10617, Taiwan\\
}

@@

% Abstract of the paper
\begin{abstract}
The \textit{Fermi Gamma-Ray Space Telescope} reveals two large bubbles in the Galaxy,\
which extend nearly symmetrically $\sim50^{\circ}$ above and below the Galactic center (GC).\
The recent discovery of giant eROSITA bubbles also shows such a symmetry about the GC, suggesting that they may originate from a single GC activity.\
Previous simulations of bubble formation that invoke active galactic nucleus (AGN) jets\
have assumed that the jets are vertical to the Galactic plane;\
however, in general there does not need to be a correlation between the jet orientation and the rotational axis of the Galactic disk.\
Using three-dimensional (3D) special relativistic hydrodynamic (SRHD) simulations that\
include cosmic rays (CRs) and thermal gas, we show that a thin interstellar medium (ISM) disk is an essential\
component for the symmetric \textit{Fermi} and eROSITA bubbles (collectively, Galactic bubbles).\
We also produce simulated gamma-ray and\
microwave spectra from the inverse Compton (IC) scattering and synchrotron,\
and compare them with the observed counterparts in order to verify\
an oblique AGN jets model. We find that\
(1) The simulated Galactic bubbles are nearly symmetric about the GC albeit\
the bipolar jets are at an angle $45^{\circ}$ with respect to\
the rotation axis of the Galatic plane (disc normal hereafter).\
(2) The edge of the eROSITA bubbles is a forward shock,\
originally driven by oblique bipolar jets emanating from the GC 12.39 million years ago,\
and significantly stretched by the stratified atmosphere afterwards.\
(3) Followed by the forward shock is a tangling contact discontinuity,\
which we interpret as the edge of the \textit{Fermi} bubbles\
composed of turbulent and high-temperature ($\sim2$ keV)\
plasma in pressure balance with the external medium.
(4) According to the microwave and leptonic gamma-ray spectra, the best-fitting\
CRe power-law index is found to be 2.4.\
The broad agreement between simulated and observed multiwavelength features\
has thus released a caveat:\
the jets shall be vertical to the disc normal.
\end{abstract}

@@

Their symmetry about the GC suggests that they originate\
from powerful energy injections from the GC, possibly related to nuclear star formation\
\citep{PhysRevLett.106.101102,Carretti2013}
or past AGN activity \citep{Guo2012,Li,Yang2017,Zhu}.

@@

 Followed by the forward shock is a turbulent and hot plasma.\
 As shown in the temperature or density slices of the fiducial run,\
 the full extent of the turbulent plasma is approximately agree with\
 that of the observed \textit{Fermi} bubbles \citep{Su2010}.
 Also, the temperature of the plasma is as high as 2 keV,\
 comparable to few keV inside the \textit{Fermi} bubbles\
 obtained by observing X-ray absorption lines through the hot\
 gaseous halo along many different sight lines in the sky \citep{Miller_2013}.\
 We also note that the turbulence within the hot plasma is in pressure balance with the external medium,\
 suggesting the outer edge of the \textit{Fermi} bubbles\
 is a turbulence rather than a smooth shock \citep{Zhang2020}.

@@

 Moreover, the close-up view (right column in Fig. \ref{fig__profile}) of profiles\
 and slices (Fig. \ref{fig__innerbubbles})\
 demonstrate that there is a sharp pressure jump on the edge of the innermost bubbles ($z=3.62$ kpc),\
 indicating that the innermost bubbles are an expanding reverse shock.\
 The high density upstream of the reverse shock requires\
 an even higher density downstream.\
 Continuing outward, the higher density must match with low density further downstream;\
 thus there exists a dense shell.
